Surrogate model for aligned-spin binary neutron star waveforms using gaussian process regression

### 2017-06-28 paper/paper.tex
Start section on EOB

%% packages
\usepackage{graphicx,psfrag}
\usepackage{mathrsfs,amsmath,amsfonts,amssymb, bm}
\usepackage{multirow}
\usepackage{comment,hyperref}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algpseudocode} % in place of algorithmic. Conflicts with revtex4 unless [H] option passed
\usepackage{setspace} % for pleasent spacing in algorithm

@@

%% ________________________________________________________
\section{Aligned-spin, dynamical tides EOB model}

--Key features of dynamical tides EOB model

@@

%% ________________________________________________________
%\bibliographystyle{revtex}     
\bibliography{paper}

@@

>>>>

%% packages
\usepackage{graphicx,psfrag}
\usepackage{mathrsfs,amsmath,amsfonts,amssymb, bm}
\usepackage{multirow}
\usepackage{comment,hyperref}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algpseudocode} % in place of algorithmic. Conflicts with revtex4 unless [H] option passed
\usepackage{setspace} % for pleasent spacing in algorithm
\usepackage{times}

@@

%% ________________________________________________________
\section{Aligned-spin, dynamical tides EOB model}
Ever since its inception~\cite{Buonanno:1998gg}, the effective-one-body (EOB) approach to the general-relativistic 2-body problem has proven successful in modeling the dynamics and GW emission of binaries of compact objects. State-of-the-art aligned-spin EOB models~\cite{Bohe:2016gbl,Nagar:2017jdw} can accurately match numerical-relativity (NR) simulations of spinning, nonprecessing binary black holes (BBHs) for mass ratios up to 8 and spin magnitudes up to 0.85 for unequal-mass (up to almost extremal for equal-mass) binaries. The EOB framework can also accommodate precessing-spin BBHs, showing good agreement to mildly precessing NR simulations~\cite{Babak:2016tgq}. This research program has been extended to accommodate matter effects~\cite{}, which become relevant XXX

References~\cite{Hinderer:2016eia,Steinhoff:2016rfi} built upon the aligned-spin BBH EOB model of Ref.~\cite{Taracchini:2013rva} and proposed a way to include the effect of dynamical tides. Neutron stars that are part of a compact-object binary will deform in the tidal field generated by the companion --- be it a black hole or another neutron star. Because of binary motion, the forcing tidal field varies at the orbital frequency. Thus, in the late stages of the inspiral, the characteristic $f$-mode frequency of the neutron star can be dynamically approached, resulting in a resonant excitation of such mode. The net effect is an amplification of tidal effects as compared to the adiabatic limit, which assumes that the $f$-mode frequency is much larger than the frequency of the forcing tidal field. Dynamical tidal effects are implemented in the EOB model through a modification of the radial potential $\Delta_u$, which is the $tt$-component of the metric of the effective spacetime. Here we adopt the tidally-augmented expression of $\Delta_u$ that is discussed in Appendix~A of Ref.~\cite{Steinhoff:2016rfi}, namely
\begin{equation}
\Delta_u = \Delta_u^{\textrm{pm}} + \Delta_u^{\textrm{DT}}\,,
\end{equation}
where $\Delta_u^{\textrm{pm}}$ is the 4PN-accurate point-mass term of the underlying BBH EOB model of Ref.~\cite{Bohe:2016gbl} (see Eq.~(2.2) therein) and $\Delta_u^{\textrm{DT}}$ is the contribution due to dynamical tides. We consider the general case of a binary that may consist of two neutron stars; if either component of the binary is a black hole, then vanishing values of the tidal polarizabilities are employed in the formulas below. Let $A$ and $B$ be the body labels. Let $M^{A,B}$ be the mass of either body and $M=M^A+M^B$ be the total mass of the binary.  Explicitly, including quadrupolar and octupolar tidal effects, we have
\begin{widetext}
\begin{align}
 \Delta_u^{\textrm{DT}} =&  - 3\,\Lambda_{2,\textrm{dyn}}^{A}(u)\,u^6\frac{X_B}{X_A} \left[ 1 + \frac{5}{2} X_Au + \left(3 + \frac{1}{8}X_A + \frac{337}{28} X_A^2\right)u^2 \right]\nonumber\\
 &- 15\,\Lambda_{3,\textrm{dyn}}^{A}(u)\,u^8 \frac{X_B}{X_A}\left[1 + \left(-2 + \frac{15}{2}X_A\right)u + \left(\frac{8}{3} - \frac{311}{24}X_A + \frac{110}{3}X_A^2\right) u^2\right] + (A \leftrightarrow B)\,,
\end{align}
\end{widetext}
where $X_{A,B}=M^{A,B}/M$, $u=1/r$ is the inverse of the ($M$-rescaled) EOB radial coordinate, $\Lambda_{2,\textrm{dyn}}^{A,B}(u)$ are the dimensionless quadrupolar dynamical tidal polarizabilities, and $\Lambda_{3,\textrm{dyn}}^{A,B}(u)$ are the dimensionless octupolar dynamical tidal polarizabilities. Within the dynamical tides model, the tidal polarizabilities are not constant, but rather depend on the orbital separation and on the values of the $f$-mode frequencies. In particular, the dimensionless dynamical polarizabilities read
\begin{equation}
\Lambda_{\ell,\textrm{dyn}}^{A,B}(u)=\frac{2}{(2\ell-1)!!}k^{A,B}_{\ell}\hat{k}^{A,B}_{\ell\,\textrm{dyn}}(u;\omega_{0\ell}^{A,B})\left(\frac{R_{A,B}}{M}\right)^{2\ell+1}\,,
\end{equation}
where $k^{A,B}_{\ell}$ is the tidal Love number and $R_{A,B}$ is the neutron star radius. Here $\hat{k}^{A,B}_{\ell\,\textrm{dyn}}(u;\omega_{0\ell}^{A,B})$ is the separation-dependent, dimensionless enhancement factor given in Eq.~(11) of Ref.~\cite{Dietrich:2017feu}, which depends on the value of the $f$-mode angular frequency of the $2^\ell$-th multipole, $\omega_{0\ell}^{A,B}$.

Tidal effects also enter the dissipative part of the model. In particular, the point-mass $\ell=2,3$ inspiral-plunge waveform modes are corrected by tidal terms
\begin{equation}
h_{\ell m}^{\textrm{insp-plunge}} = h_{\ell m}^{\textrm{insp-plunge, pm}} + h_{\ell m}^{\textrm{tidal}}\,,
\end{equation} 
where $h_{\ell m}^{\textrm{insp-plunge, pm}}$ is discussed in Section~II.B of Ref.~\cite{Bohe:2016gbl} and $h_{\ell m}^{\textrm{tidal}}$ is given by Eqs. (A14)-(A17) of Ref.~\cite{Damour:2012yf}. Following Ref.~\cite{Dietrich:2017feu}, in the computation of $h_{22}^{\textrm{tidal}}$ we include a dynamical enhancement factor that depends on the orbital separation and on the $f$-mode frequency, and is given in Eq.~(15) of Ref.~\cite{Dietrich:2017feu}.\red{[Tanja should confirm for 21]}

@@

%% ________________________________________________________
%\bibliographystyle{revtex}     
\bibliography{paper,refs}

@@

\begin{widetext}
\begin{align}
\Lambda_{2,\textrm{dyn}}^{\textrm{(A,B)}}(u) = &\frac{\lambda_{2}^{\textrm{(1,2)}}}{M^5} \left\{\frac{1}{4} + \frac{3}{4} \left(\frac{\omega^{(1,2)}_{f,2}}{2\Omega}\right)^2\left[\frac{4\Omega^2}{(\omega^{(1,2)}_{f,2})^2 - 4\Omega^2} + \frac{10}{3Q}\right.\right. \nonumber\\
&\left.\left.+\sqrt{\frac{\pi}{3}}\frac{1}{\epsilon_{1,2}}\left[(1 + 2\mathcal{S}{(y_{1,2})})\cos{\left(\frac{\pi}{2}y^2_{1,2}\right)} - (1 + 2\mathcal{C}{(y_{1,2})})\sin{\left(\frac{\pi}{2}y_{1,2}^2\right)}\right]\right]\right\}\,,\\
\Lambda_{3,\textrm{eff}}^{\textrm{(1,2)}} (u)= &\frac{\lambda_{3}^{\textrm{(1,2)}}}{M^7} \left\{\frac{3}{8} + \frac{(\omega^{(1,2)}_{f,3})^2}{\Omega^2}\left(\frac{25}{48 O_{1,2}} + \frac{5}{72}\frac{9\Omega^2}{(\omega^{(1,2)}_{f,3})^2-9\Omega^2}\right)\right. \nonumber\\
&\left.+ O_{1,2}\left[\cos{(w^2)}\left(\frac{1}{2} +\mathcal{S}(sqrt2pi*w)\right) - \sin{(w^2)}\left(\frac{1}{2} + \mathcal{C}(sqrt2pi*w)\right)\right]\right\}\,,
\end{align}
\end{widetext}
where $\lambda_{2}^{\textrm{(1,2)}}$ are the (constant) adiabatic values of the quadrupolar tidal polarizabilities, $\lambda_{3}^{\textrm{(1,2)}}$ are the (constant) adiabatic values of the octupolar tidal polarizabilities, $\omega^{(1,2)}_{f,2}$ are the quadrupolar $f$-mode frequencies, $\omega^{(1,2)}_{f,3}$ are the octupolar $f$-mode frequencies, $\mathcal{S}$ is the Fresnel sine integral, $\mathcal{C}$ is the Fresnel cosine integral, $\epsilon_{1,2} = 2^{19/3}\nu (M\omega^{(1,2)}_{f,2})^{5/3}/5$,   $y_{1,2}=\sqrt{3/(\pi \epsilon)} Q_{1,2}/5$, and $Q_{1,2} = 4- 2^{1/3}(M\omega^{(1,2)}_{f,2})^{5/3}u^{-5/2}$. 
, $w_{1,2}=O_{1,2}/(4\,3^{2/3}\sqrt{10}(M\omega^{(1,2)}_{f,3})^{5/6}\sqrt{\nu})$, $O_{1,2}=5\sqrt{5\pi}/u^3(M\omega^{(1,2)}_{f,3})^{7/6}/(192\,3^{2/3}\sqrt{\nu})$.
